### 2014-05-09 learn_intro.tex
Writethru to end of Remy intro

\section{Introduction}
\label{s:intro}

Over the last 30 years, Internet congestion control has seen
considerable research interest. Starting with seminal work in the
1980s~\cite{decbit,Jacobson88,chiujain}, the Transmission Control
Protocol (TCP) has adopted a series of end-to-end algorithms to share
network resources among contending
endpoints~\cite{newreno,vegas,cubic,compound}. Another line of work
has explored the use of in-network algorithms running at bottleneck
routers to help perform this function more
efficiently~\cite{Floyd93,ecn,BLUE,AVQ,choke,CoDel,xcp}.

As the Internet grows and evolves, it appears likely that new network
protocols will continue to be developed to accommodate new subnetwork
behaviors and shifting application workloads and goals. Some of these
may be intended for specialized environments---e.g., inside a
centrally-managed datacenter---while some will be for broad use across
the wide-area Internet, or over cellular network paths.

In practice, however, it is challenging to guarantee that a
distributed system's performance in well-characterized testing
scenarios will extend to real networks, which inevitably differ from
those envisioned in development and will continue to evolve over time. This
uncertain generalizability presents an obstacle to any new protocol's
deployment.

In this paper, we formalize the design process for generating
end-to-end congestion-control protocols to quantify: \emph{how easy is
  it to ``learn'' a network protocol to achieve desired goals, given a
  necessarily imperfect model of the networks where it will ultimately
  be deployed?}

@@

\subsection{Tractable attempts at optimal (Tao) protocols}

Each of the above areas of inquiry is about the effect of a protocol
designer's imperfect understanding of the future network that a
decentralized congestion-control protocol will ultimately be run over.

In principle, we would quantify such an effect by evaluating, on the
actual network, the performance of the ``best possible''
congestion-control protocol designed for the imperfect network
\emph{model}, and comparing that with the best-possible protocol for
the actual network itself.

In practice, however, there is no tractable way to calculate the
optimal congestion-control protocol for a given network. Instead,
throughout this study we use the Remy automatic protocol-design
tool~\cite{remy} as a proxy for the optimal solution.

We refer to these automatically-generated congestion-control protocols
as ``tractable attempts at optimal'' (Tao) end-to-end
congestion-control protocols for a given network.  Tao protocols
represent a practically realizable tilt at developing an optimal
protocol for an imperfect model of a real network.

Constructing a Tao for a complex network model requires searching a
huge space, an intensive task even using Remy's search
optimizations. In most cases, the protocol stopped improving within a
CPU-year of processing time (five days on our 80-core machine), though
there were a few cases where improvements continued to occur.

%some cases in our investigation, the protocol continued
%to improve even after the search had consumed more than a CPU-year of
%processing time, or more than five days on an 80-core computer.

\begin{figure}[t!]
\includegraphics[width=\columnwidth]{figures/plots-10x-new-multilink/multilink-crossing-manual.pdf}
\caption{How well do endpoints need to understand the network around
  them? To assess this, we measure the throughput of four
  congestion-control schemes across a simulated two-hop network path,
  as the rate of each link is swept between 10 and 100~Mbps, with
  75~ms of delay per hop. A tractable attempt at optimal (Tao)
  congestion control for a simplified one-bottleneck model of the
  network performs only a little worse than a protocol designed with
  knowledge of the network's full two-bottleneck structure.}
\label{f:multihop}
\end{figure}

We emphasize that there can be no assurance that the Tao actually
comes close to the optimal congestion-control protocol, except to the
extent that it approaches upper bounds on performance, such as the
ideal fair allocation of network resources. But empirically, such
protocols perform well. Figure~\ref{f:multihop} shows the end-to-end
throughput of connections traversing two bottleneck links for
different protocols, including two Tao protocols. In each experiment,
there were three flows: one that traversed two hops (the measured
connection), a cross-traffic flow that traversed only the first
bottleneck, and another cross-traffic flow that traversed only the
second bottleneck.

%of two Tao protocols, running over a simulated two-hop network path
%and contending with cross traffic that runs over the first or
%second link only.

As the rate of the first and second link are swept between 10 and 100
Mbps, the Tao that was designed for the actual network achieves close
to the ideal proportionally-fair throughput allocation across the two
links. This ideal serves as an upper bound on performance.

A Tao that was designed for a simplified model of the network with
only one hop also performs adequately, but a little worse than the
``full-complexity'' Tao. However, both computer-generated protocols
come closer to the ideal than contemporary protocols in wide use, TCP
Cubic~\cite{cubic}, the default TCP in Linux, and Cubic running over
stochastic fair queueing and CoDel~\cite{CoDel}, an active queue
management (AQM) scheme that runs at both bottlenecks.

%Prior work on computer-generated congestion-control
%protocols~\cite{remy} also found that such algorithms generally
%outperform the breadth of human-designed protocols on the network for
%which they were designed, suggesting they are a fair proxy for the
%optimal congestion-control protocol given a particular network model.

\subsection{Summary of results}

Here are the principal findings of this study:

@@

\noindent 
\textbf{Modeling a two-bottleneck network as a single bottleneck hurt
  performance only mildly.}  On the two-hop network of
Figure~\ref{f:multihop}, on average we find that a protocol designed
for a simplified, single-bottleneck model of the network
underperformed by {\bf only 17\%} a protocol that was designed with
full knowledge of the network's two-bottleneck structure. The
simplified protocol outperformed TCP Cubic by a factor of $7.2\times$
on average, outperformed Cubic-over-sfqCoDel by $2.75\times$ on
average.

Thus, in this example, full knowledge of the network topology during
the design process was not crucial.

@@

\noindent 
\textbf{Only weak evidence of a tradeoff between
  operating range and performance.} We created a Tao protocol designed
for a range of networks whose link rates spanned a thousand-fold range
between 1~Mbps and 1000~Mbps, as well as three other protocols that
were more narrowly-targeted at hundred-fold, ten-fold, and two-fold
ranges of link rate (Figure~\ref{fig:breadth}).

The ``thousand-fold'' Tao achieved close to the peak performance of
the ``two-fold'' Tao. Between link rates of 22--44~Mbps, the
``thousand-fold'' Tao achieved \textbf{within 3\% of the throughput}
of the ``two-fold'' protocol that was designed for this exact
range. However, the queueing delay of the ``thousand-fold'' protocol
was \textbf{46\% higher}, suggesting some benefit from more focused
operating conditions. It also takes a lot longer to compute (offline)
a ``thousand-fold'' Tao compared to a two-fold Tao; in run-time
operation, though, the computational cost of the two algorithms is
similar.

Over the full range of 1~Mbps to 1000~Mbps, the ``thousand-fold'' Tao
protocol matched or outperformed TCP Cubic and Cubic-over-sfqCoDel
over the entire range (Figure~\ref{fig:breadth}).

The results suggest that highly optimized protocols may be able to
perform adequately over a broad range of actual networks. Broad
operating range had only a weak effect on peak performance, suggesting
that ``one size fits all'' congestion-control protocols that perform
considerably better than TCP---as well as TCP plus in-network
assistance, in the case of sfqCoDel---may be feasible.

@@

When contending only with other endpoints of the same kind, the
``TCP-naive'' protocol achieved \textbf{55\% less queueing delay} than
the TCP-aware protocol. In other words, ``TCP-awareness'' has a cost
measured in lost performance when TCP cross-traffic is not present
(Figure~\ref{fig:tcpaware}).

But when contending against TCP, the ``TCP-naive'' protocol was
squeezed out by the more-aggressive cross-traffic. By contrast, the
``TCP-aware'' protocol achieved \textbf{36\% more throughput} and
\textbf{37\% less queueing delay} than the ``TCP-naive'' protocol, and
claimed its fair share of the link capacity from TCP.

>>>>

%\section{Introduction}
\label{s:intro}

The previous chapter described the development of Remy, our automated
protocol-design tool. Remy generates congestion-control schemes from
first principles, given an uncertain model of the underlying network
and a stated application objective to pursue.

In practice, the designer of a congestion-control scheme for the
wide-area Internet cannot expect to supply a protocol-design tool with
a perfectly faithful model of the actual Internet. An actual model
will necessarily be simplified and imperfect. Nonetheless, designers
would naturally like to develop protocols that are broadly useful on
real networks.

In this chapter, we use Remy to quantify how much tolerance the
protocol designer can expect, or in other words: \emph{How easy is it
  to ``learn'' a network protocol to achieve desired goals, given a
  necessarily imperfect model of the networks where it will ultimately
  be deployed?}

@@

%\subsection*{RemyCCs as proxies for the best known algorithm}

Each of the above areas of inquiry is about the effect of a protocol
designer's imperfect understanding of the future network that a
decentralized congestion-control protocol will ultimately be run
over. In principle, we could quantify such an effect by evaluating, on
the actual network, the performance of the ``best possible''
congestion-control protocol designed for the imperfect network
\emph{model}, and comparing that with the best-possible protocol for
the actual network itself.

In practice, however, there is no tractable way to calculate the
optimal congestion-control protocol for a given network. Instead,
throughout this chapter we use Remy to construct achievability
results---RemyCCs---that stand in for the ``best possible'' average
performance across the range of network parameters supplied by the
designer.

We emphasize that there can be no assurance that Remy's algorithms
actually comes close to the optimal congestion-control protocols,
except to the extent that they approache analytical upper bounds on
performance and outperform existing schemes across the same networks.

\section{Summary of results}

The principal findings of these experiments are as follows:

@@

\noindent 
\textbf{Modeling a two-bottleneck network as a single bottleneck hurt
  performance only mildly.}  On the two-hop network of
Figure~\ref{fig:two-link}, on average we find that a protocol designed
for a simplified, single-bottleneck model of the network
underperformed by {\bf 17\%} a protocol that was designed with full
knowledge of the network's two-bottleneck structure
(Figure~\ref{f:multihop}). The simplified protocol outperformed TCP
Cubic-over-sfqCoDel by $2.75\times$ on average. In this example, full
knowledge of the network topology during the design process was not
crucial.

@@

\noindent 
\textbf{Only weak evidence of a tradeoff between
  operating range and performance.} We created a RemyCC designed
for a range of networks whose link rates spanned a thousand-fold range
between 1~Mbps and 1000~Mbps, as well as three other protocols that
were more narrowly-targeted at hundred-fold, ten-fold, and two-fold
ranges of link rate (Figure~\ref{fig:breadth}).

The ``thousand-fold'' RemyCC achieved close to the peak performance of
the ``two-fold'' RemyCC. Between link rates of 22--44~Mbps, the
``thousand-fold'' RemyCC achieved \textbf{within 3\% of the throughput}
of the ``two-fold'' protocol that was designed for this exact
range. However, the queueing delay of the ``thousand-fold'' protocol
was \textbf{46\% higher}, suggesting some benefit from more focused
operating conditions. It also takes Remy longer to optimize (offline)
a ``thousand-fold'' RemyCC compared with a two-fold RemyCC. In run-time
operation, the computational cost of the two algorithms is
similar.

Over the full range of 1~Mbps to 1000~Mbps, the ``thousand-fold''
RemyCC protocol matched or outperformed TCP Cubic and
Cubic-over-sfqCoDel over the entire range
(Figure~\ref{fig:breadth}). This result suggests that it may be
possible to construct ``one size fits all'' end-to-end
congestion-control protocols that perform considerably better than
TCP, even when TCP is assisted by in-network algorithms.

@@

When contending only with other endpoints of the same kind, the
``TCP-naive'' protocol achieved \textbf{55\% less queueing delay} than
the TCP-aware protocol. In other words, ``TCP-awareness'' has a cost
measured in lost performance when TCP cross-traffic is absent
(Figure~\ref{fig:tcpaware}).

But when contending against TCP, the ``TCP-naive'' protocol was
squeezed out by the more-aggressive cross-traffic. By contrast, the
``TCP-aware'' protocol achieved \textbf{36\% more throughput} and
\textbf{37\% less queueing delay} than the ``TCP-naive'' protocol,
and reached an equitable division of the available capacity with TCP.

### 2014-05-09 learn_operational.tex
Writethru to end of Remy intro

\begin{figure}
\centering
%\begin{subfigure}[b]{0.49\textwidth}
%\includegraphics[width=\textwidth]{tradeoff-tpt.pdf}
%\end{subfigure}
%\begin{subfigure}[b]{0.33\textwidth}
%\includegraphics[width=\textwidth]{tradeoff-delay.pdf}
%\end{subfigure}
%\begin{subfigure}[b]{0.49\textwidth}
%\includegraphics[width=\textwidth]{tradeoff-util.pdf}
%\end{subfigure}
\includegraphics[width=\columnwidth]{figures/parametric-oprange/oprange-manual.pdf}
\caption{Evidence of a weak tradeoff between operating range of a
  congestion-control protocol and performance. The Tao protocols
  designed with more specific network models (Tao-2x and Tao-10x)
  performed modestly better---within their design ranges---than protocols
  designed for a broader range of networks (Tao-100x and Tao-1000x),
  at a cost of deterioration when the actual network did not fall
  within the training scenarios. The four Tao protocols outperformed
  Cubic and Cubic-over-sfqCoDel over their design ranges.}
\label{fig:breadth}
\end{figure}

@@

@@

The results suggest that a congestion-control protocol designed for a
simplified model of the network's structure will experience a small,
but quantifiable, penalty to its performance, indicated on the figure
as the gap between the two Tao protocols. Results from Cubic and
Cubic-over-sfqCoDel are shown for comparison.

These results give some rigorous support to a protocol designer's
intuition that understanding the true network in all its complexity
may not be crucial.

>>>>

\begin{figure}
\centering
%\begin{subfigure}[b]{0.49\textwidth}
%\includegraphics[width=\textwidth]{tradeoff-tpt.pdf}
%\end{subfigure}
%\begin{subfigure}[b]{0.33\textwidth}
%\includegraphics[width=\textwidth]{tradeoff-delay.pdf}
%\end{subfigure}
%\begin{subfigure}[b]{0.49\textwidth}
%\includegraphics[width=\textwidth]{tradeoff-util.pdf}
%\end{subfigure}
\includegraphics[width=\columnwidth]{oprange-manual.pdf}
\caption{Evidence of a weak tradeoff between operating range of a
  congestion-control protocol and performance. The RemyCC protocols
  designed with more specific network models (RemyCC-2x and RemyCC-10x)
  performed modestly better---within their design ranges---than protocols
  designed for a broader range of networks (RemyCC-100x and RemyCC-1000x),
  at a cost of deterioration when the actual network did not fall
  within the training scenarios. The four RemyCC protocols outperformed
  Cubic and Cubic-over-sfqCoDel over their respective design ranges.}
\label{fig:breadth}
\end{figure}

@@

\begin{figure}[t!]
\includegraphics[width=\columnwidth]{multilink-all.pdf}
\caption{How well do endpoints need to understand the network around
  them? To assess this, we measure the throughput of four
  congestion-control schemes across a simulated two-hop network path,
  as the rate of each link is swept between 10 and 100~Mbps, with
  75~ms of delay per hop. A RemyCC designed for a simplified
  one-bottleneck model of the network performs 17\% worse, on average,
  than a RemyCC designed with knowledge of the network's true
  two-bottleneck structure. Both RemyCCs outperform TCP Cubic assisted
  by per-flow queueing and CoDel at the bottlenecks.}
\label{f:multihop}
\end{figure}

@@

Figure~\ref{f:multihop} shows the end-to-end throughput of connections
traversing two bottleneck links for different protocols, including two
RemyCCs. In each experiment, there were three flows: one that
traversed two hops (the measured connection), a cross-traffic flow
that traversed only the first bottleneck, and another cross-traffic
flow that traversed only the second bottleneck.

%of two Tao protocols, running over a simulated two-hop network path
%and contending with cross traffic that runs over the first or
%second link only.

As the rate of the first and second link are swept between 10 and 100
Mbps, the RemyCC that was designed for the actual network achieves
close to the ideal proportionally-fair throughput allocation across
the two links. This ideal serves as an upper bound on performance.

A RemyCC that was designed for a simplified model of the network with
only one hop also performs adequately, but a little worse than the
``full-complexity'' RemyCC. However, both computer-generated protocols
come closer to the ideal than contemporary protocols in wide use, TCP
Cubic~\cite{cubic}, the default TCP in Linux, running over per-flow
queueing and CoDel~\cite{CoDel}, an active queue management (AQM)
scheme that runs at both bottlenecks.

The results suggest that a congestion-control protocol designed for a
simplified model of the network's structure will experience a small,
but quantifiable, penalty to its performance.
